### 2016-09-25 reviews/reviewer3-response.md

# Response to reviewer 3

We appreciate the comments given here, and addressing them has improved the manuscript significantly.

>1. For convergence study, Fig 5 gives the results for different choice of time step. however, for SST model, the Cp do not convergent with the time step. On the other hand, the choice of time step is always coupled with the choice of space step. Generally speaking, the time step cannot be too small for a certain space discretion condition. Therefore, It is suggested that the author provide this relationship between space and time discretion.

>2. No the convergence study for 3d model in this paper.

>3. In Performance Prediction session, the author gives a figure(Fig.6 ) to show the comparison of 3d , 2d and experimental data. But fig.6 only gives the results for one case for different methods. The 2d method seems over-predict the Cp and Cd. more results of different cases should be given to verify this idea.

>4. Shaft and Arm/Strut effects are important to the turbine, and have been discussed by previous publications by using RANS such as
>
Behrouzi F, Malik A M A, Nakisa M, et al. Arm Effect on Performance of Vertical Axis Current Turbine Using RANS Simulation. Applied Mechanics & Materials, 2014, 554(554):531-535.
>
and by using experimental test such as
Li Y, Calisal S M. Three-dimensional effects and arm effects on modeling a vertical axis tidal current turbine. Renewable Energy, 2010, 35(10):2325-2334.
>
However, the authors have not well explained it in the manuscript. Please extend this part.