### 2016-10-02 reviews/reviewer2-response.md
Discuss effect of slip BC

# Response to reviewer 2

We thank the reviewer for his or her thoughtful comments, which have helped us
improve the paper.

>In general more work needs to be performed on
• Discussion of mesh factors that may affect the simulation results

>• Nomenclature, and the lack of definition of many of the variables will cause confusion for many readers

>• Improvement of figures with nomenclature, labels, axis etc.

>• More explanation of why the various transport properties are investigated and their applicability

>Additionally the title could be better, maybe add a Computational Fluid Dynamics or simulation somewhere?

## Responses to PDF comments

>Effect of this [slip boundary condition] on the results?

It didn't seem to matter much. The mean velocity fields appear to be nearly
symmetrical about the $x$--$y$ plane, but nonetheless, the slip BC is at least
closer to how a free surface would behave than a no slip BC.