### 2016-10-02 reviews/reviewer2-response.md
Comment on near-blade mesh resolution

# Response to reviewer 2

We thank the reviewer for his or her thoughtful comments, which have helped us
improve the paper.

>In general more work needs to be performed on
• Discussion of mesh factors that may affect the simulation results

>• Nomenclature, and the lack of definition of many of the variables will cause confusion for many readers

>• Improvement of figures with nomenclature, labels, axis etc.

>• More explanation of why the various transport properties are investigated and their applicability

>Additionally the title could be better, maybe add a Computational Fluid Dynamics or simulation somewhere?

## Responses to PDF comments

>Effect of this [slip boundary condition] on the results?

It didn't seem to matter much. The mean velocity fields appear to be nearly
symmetrical about the $x$--$y$ plane, but nonetheless, the slip BC is at least
closer to how a free surface would behave than a no slip BC.

>Is this mesh [Figure 4] too coarse for adequate resolution of the flow?

It could be. For our mesh dependence study we held the topology fixed and scaled
the mesh proportionally. The mesh should be fine enough to resolve the boundary
layer since $y^+ \sim 1$, but it is a fair point that maybe separated flow
structures may be dissipated in an exaggerated way. However, exploring the mesh
topology parameter space would have been prohibitive.

>Why blank? [viscous transport in Figure 12]

These terms aren't zero, but are extremely small.