### 2016-10-09 paper.tex
Discuss strut drag modeling effects

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for
% %  REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with
% the REVTEX 4.1 document class and the AIP substyles.
%
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}
\usepackage[draft]{todonotes}
\usepackage{subcaption}

\draft % marks overfull lines with a black rule on the right
\usepackage{graphicx}
\graphicspath{{figures/}}

\begin{document}

% Use the \preprint command to place your local institutional report number
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
\preprint{Preprint submitted to Journal of Renewable and Sustainable Energy
19 March 2016}

\title{Modeling the near-wake of a vertical-axis cross-flow turbine with 2-D and
    3-D RANS}

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
% Explanatory text should go in the []'s,
% actual e-mail address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro for the type of information

% \affiliation command applies to all authors since the last \affiliation command.
% The \affiliation command should follow the other information.

\author{Peter Bachant}
\email[]{pete.bachant@gmail.com}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{Center for Ocean Renewable Energy, University of New Hampshire,
Durham, NH}

\author{Martin Wosnik}
%\email[]{}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{Center for Ocean Renewable Energy, University of New Hampshire,
Durham, NH}

% Collaboration name, if desired (requires use of superscriptaddress option in
% \documentclass).
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
    The near-wake of a vertical-axis cross-flow turbine (CFT) was modeled
    numerically via blade-resolved $k$--$\omega$ SST and Spalart--Allmaras RANS
    models in two and three dimensions. Results for each case were compared with
    experimental measurements of the turbine shaft power, overall streamwise
    rotor drag, mean velocity, turbulence kinetic energy, and momentum transport
    terms in the near-wake at one diameter downstream. It was shown that 2-D
    simulations overpredict turbine loading and do not resolve mean vertical
    momentum transport, which plays an important role in the near-wake's
    momentum balance. The 3-D simulations fared better at predicting
    performance, with the Spalart--Allmaras model predictions being closest to
    the experiments. The SST model more accurately predicted the turbulence
    kinetic energy while the Spalart--Allmaras model more closely matched the
    momentum transport terms in the near-wake. These results show the potential
    of blade-resolved RANS as a design tool and a way to ``extrapolate''
    experimental flow field measurements.
\end{abstract}

\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs
% \listoftodos

\section{Introduction}

In the pursuit of a fully sustainable energy profile, cross-flow (often
vertical-axis) turbines (CFTs) can still play a role. Despite extensive research
and development in the 1970s through the early 1990s by groups like Sandia
National Labs in the US \cite{Sutherland2012} and the National Research Council
in Canada \cite{Para2002}, vertical-axis cross-flow Darrieus turbines were all
but abandoned for large scale commercial terrestrial wind power. Today however,
with the development of marine hydrokinetic (MHK) energy devices, CFTs are at
the forefront, with the first grid-connected commercial device being a
horizontal-axis CFT manufactured and installed by ORPC in Cobscook Bay, Maine
\cite{ORPC2012}. CFTs are also being examined for small-scale onshore wind in
urban areas \cite{Lott2015}, offshore floating wind farms \cite{Paulsen2011,
Sandia2012}, and onshore wind farms where power output per unit land area is of
interest \cite{Dabiri2011}.

Despite the development of many simple engineering models based on blade element
momentum or vortex methods, it remains difficult to predict the performance of
cross-flow turbines when solidity, or blade chord-to-radius ratio is high---a
common characteristic of smaller rotors, and those designed for marine
hydrokinetic (MHK) applications, since water's higher density increases rotor
loading by roughly an order of magnitude for the same frontal
area~\cite{Bachant2016-thesis}. With computing power becoming evermore available
and affordable, CFD based on the Reynolds-averaged Navier--Stokes (RANS)
equations has become an attractive method for predicting the performance of
cross-flow turbines. Since a direct numerical simulation (DNS) of all scales at
realistic Reynolds numbers is not currently feasible, turbulence be modeled.
However, with an appropriate turbulence model, blade-resolved RANS presents a
more physically realistic first principles based approach versus simpler models
based on momentum theory or potential flow. However, CFD can be computationally
expensive when done in three dimensions, but may be necessary to predict
performance for low aspect ratio or variable radius rotors, and to predict the
unique flow field produced by CFTs, and therefore wake recovery and interaction,
which is discussed in detail later.

There are many examples in the literature of 2-D cross-flow turbine simulations
with widely varying results. Balduzzi et al.~\cite{Balduzzi2016} provides a
summary of recent efforts and an attempt to standardize a methodology for using
Reynolds-averaged Navier--Stokes (RANS) to correctly predict performance of a
2-D CFT. Howell et al.~\cite{Howell2010}, performed both 2-D and 3-D simulations
of a high solidity cross-flow turbine using a $k$--$\epsilon$ renormalization
group (RNG) turbulence model. The results from the 2-D simulations overpredicted
power coefficient
\begin{equation}
    C_P = \frac{P}{\frac{1}{2} \rho A U_\infty^3},
    \label{eq:cp}
\end{equation}
where $P$ is the mechanical shaft power, $\rho$ is the fluid density, $A$ is the
rotor frontal area, and $U_\infty$ is the free stream velocity. This was likely
because the rotor had a low aspect ratio (or ratio of height to rotor diameter
$H/D$), while the 3-D case matched well with wind tunnel measurements near the
tip speed ratio of maximum power. In general, 3-D simulations are less common,
but have begun to appear more frequently recently---a testament to the progress
towards higher computing power.

An overview of 3-D blade-resolved cross-flow turbine simulations reported in the
literature is presented in Table~\ref{tab:cfd-refs}. The $k$--$\omega$ SST RANS
turbulence model is shown to be a popular choice due to its success in
predicting flows with adverse pressure gradients and
separation~\cite{Menter2003}. Higher fidelity methods that resolve the large
scales of turbulence, such as large eddy and detached eddy simulation have also
been used, though the high computational expense of these techniques may
preclude their use in industrial applications. Note that for all the studies
listed, model validation was performed for either performance or wake
predictions---not both---and in some case omitted entirely. In general, the 3-D
RANS results cannot replicate a turbine's entire performance envelope, i.e.,
$C_P$--$\lambda$ curve, but are successful for some tip speed ratios.

\begin{table}
    \centering
    \begin{tabular}{c|c|c|c}
        Author & Turbulence modeling & Perf. val. & Wake val. \\
        \hline
        Alaimo et al.~\cite{Alaimo2015} & $k$--$\epsilon$ RANS & N/A & N/A \\
        Marsh et al.~\cite{Marsh2015} & SST RANS & Reference~\cite{Rawlings2008} & N/A \\
        Orlandi et al.~\cite{Orlandi2015} & SST RANS & References~\cite{Akins1989,Mertens2003} & N/A \\
        Lam \& Peng~\cite{Lam2016} & SST RANS \& IDDES & N/A & Reference~\cite{Tescione2014} \\
        Nini et al.~\cite{Nini2014} & Spalart--Allmaras RANS & N/A & Reference~\cite{Battisti2011} \\
        Boudreau \& Dumas~\cite{Boudreau2015} & Spalart--Allmaras DDES & N/A & N/A \\
        Li et al.~\cite{Li2013} & SST RANS \& Smagorinsky--Lilly LES & Reference~\cite{McLaren2011} & N/A \\
        Howell et al.~\cite{Howell2010} & $k$--$\epsilon$ RNG RANS & Reference~\cite{Howell2010} & N/A
    \end{tabular}

    \caption{Selected 3-D blade-resolved cross-flow turbine simulations reported
        in the literature, turbulence modeling employed, and performance and/or wake
        studies used for validation. Note the Li et al. study used periodic boundary
        conditions and is technically considered 2.5-D.}

    \label{tab:cfd-refs}
\end{table}

Modeling the boundary layer flows on cross-flow turbine blades is essential to
predicting the blade loading. This flow condition presents a challenge since the
blades encounter a constantly changing relative velocity and angle of
attack---which often exceeds static stall values and causes dynamic stall.
Furthermore, the ability to predict the occurrence and interdependence of
boundary layer transition to turbulence and separation can have dramatic
influence on the blade loading and therefore the predicted turbine power output.
These challenges present significant obstacles to the prospect of using CFD to
replace wind tunnel or tank testing of physical models.

In this study we set out to model the performance and near-wake of the high
solidity University of New Hampshire Reference Vertical-Axis Turbine (UNH-RVAT)
using the Reynolds-averaged Navier--Stokes equations using the open-source
finite volume CFD package OpenFOAM, version 2.3.x. Though studies in the
literature generally focus on predicting the turbine loading and the local blade
boundary layer, we seek to model both this and the larger scale flow produced by
the rotor, i.e., the near-wake, which is of interest for this particular turbine
since it has been shown experimentally that the near-wake's momentum and energy
transport processes are dominated by vertical advection \cite{Bachant2015-JoT}.
It logically follows that a 2-D simulation, which omits the vertical dimension,
would not correctly predict wake recovery and turbine--turbine interaction.
However, it is of interest to determine how wrong a 2-D model may be, since the
lower computational cost of 2-D simulations is attractive.

We sought to validate 2-D and 3-D blade-resolved RANS models against the
UNH-RVAT mechanical power and near-wake measurements taken at the experimentally
determined optimal tip speed ratio $\lambda_0 = 1.9$. If the blade loading and
velocity in the near-wake match well enough, the flow field can potentially be
inspected in greater detail, i.e., we may be able to make observations of
quantities not measured experimentally, e.g., pressure, and gain greater insight
into where the dominant flow structures originate. This will ultimately help
develop and evaluate low-order wake generator models for use in turbine array
modeling. In summary, the questions we hope to answer here are:

\begin{enumerate}
    \item Can 2-D RANS be used for individual turbine and/or array design?

    \item How accurately can 3-D RANS predict performance?

    \item Do the flow fields predicted by 3-D RANS match the experimental wake
    measurements well enough to provide insight for developing new low-order
    wake generators to represent CFTs?

    \item Does 3-D RANS realize the correct proportions of wake recovery
    mechanisms, i.e., are the 3-D blade-resolved results a good reference case
    or ``target'' for those a low-order model should produce?
\end{enumerate}

\section{Numerical setup}

In this study the UNH-RVAT baseline tow tank experiment was simulated using a
mean rotor tip speed ratio $\lambda=1.9$, which corresponds to the maximum
measured power coefficient; cf. Figure~\ref{fig:exp-perf}. The tow speed or free
stream velocity $U_\infty=1.0$ m/s gives a turbine diameter Reynolds number
$Re_D \approx 10^6$, which corresponds to the $Re$-independent state for both
performance and near-wake characteristics, as determined from previous
experimental measurements \cite{Bachant2014, Bachant2016-Energies}.

The turbine's CAD geometry was prepared or ``cleaned'' for CFD by removing
details determined to be unnecessary, e.g., screw heads and axial shaft grooves,
which would complicate the meshing process and contribute very little to the
overall loading or flow modulation. A drawing of the physically and numerically
modeled geometries is presented in Figure~\ref{fig:cfd-cad}, and geometric
parameters are listed in Table~\ref{tab:turb-geom}.

\begin{figure}
    \centering

    \includegraphics[width=0.9\textwidth]{figures/exp_perf}

    \caption{Mean rotor power (left) and drag (right) coefficient curves from
        the tow tank experiments\cite{Bachant2016-RVAT-Re-dep}.}

    \label{fig:exp-perf}
\end{figure}

\begin{figure}
    \centering

    \includegraphics[clip, trim=0 1in 0 1in, width=0.8\textwidth]{CAD}

    \caption{CAD drawings of the UNH-RVAT cross-flow turbine as designed (left)
        and cleaned for simulation (right).}

    \label{fig:cfd-cad}
\end{figure}

\begin{table}
    \centering
    \begin{tabular}{r|l}
        Parameter & Value \\
        \hline
        Diameter $D$ & 1.0 m \\
        Height or blade span $H$ & 1.0 m \\
        Blade chord length $c$ & 0.14 m \\
        Blade profile & NACA 0020 \\
        Blade mounting point & $c/2$ and $H/2$ \\
        Strut chord length & 0.14 m \\
        Strut profile & NACA 0020
    \end{tabular}

    \caption{Geometric parameters for the UNH-RVAT rotor modeled in this study.}

    \label{tab:turb-geom}
\end{table}

To close the RANS equations, two different turbulence models were used: Menter's
$k$--$\omega$ SST \cite{Menter1994} and the Spalart--Allmaras (SA) one equation
model \cite{Spalart1992}. Both closures use the eddy-viscosity approach---the SA
employing a single additional scalar transport equation for an
eddy-viscosity-like quantity $\tilde{\nu}$ and the SST solving equations for the
turbulence kinetic energy $k$ and specific dissipation $\omega$. The SST model
was chosen due to its prominence in the literature for simulating separating
flows, which we assumed to be present in the current problem in the form of
dynamic stall~\cite{Para2002}. The SA model was shown by Ferreira et
al.~\cite{Ferreira2007} to match experimental particle image velocimetry (PIV)
results for a CFT in dynamic stall, though this was a somewhat low Reynolds
number case ($Re_c \approx 6 \times 10^4$). Further justification for using the
SA model for this case comes from Crivellini and D'Alessandro
\cite{Crivellini2014}, where they successfully modeled the laminar separation
bubble and subsequent boundary layer transition to turbulence at Reynolds
numbers similar to those investigated here.

\subsection{Computational mesh}

The computational domain was designed to match the tow tank dimensions for
comparison with previous experiments. Thus, it was modeled as a rectangular
volume 3.66 m long, 3.66 m wide, and 2.44 m tall (for the 3-D cases), with the
turbine located 1.52 m from the inlet, centered within the tank depth with its
axis oriented vertically. The rotor geometry was located at the center of a
cylindrical sliding mesh interface, which rotated at a mean tip speed ratio
$\lambda=1.9$ with a sinusoidal oscillation at the blade passage
frequency---with and amplitude of 0.19 and the first peak at 1.4 radians---to
mimic the slight deviation from the mean tip speed ratio observed in the
experiments. The 2-D mesh overview is shown in Figure~\ref{fig:2d-br-mesh} and
the blade mesh detail is shown in Figure~\ref{fig:blade-mesh}.

Meshes were generated using OpenFOAM's blockMesh and snappyHexMesh utilities.
The mesh consisted of a background hexahedral mesh, which was refined in all
three directions by a factor of 2 in a rectangular region containing the turbine
and near-wake (0.9 m upstream, 1.3 m downstream, $\pm 0.9$ m cross-stream, and
$\pm 0.8$ m vertically). Cells adjacent to the turbine shaft and struts were
refined by a factor of 4, while cells adjacent to the blades were refined by a
factor of 6. To capture the boundary layer, i.e., where flow physics most affect
rotor torque---20 layers were added next to the blades with an expansion ratio
of 1.2. Overall mesh refinement is controlled by a single parameter---the number
of cells in the streamwise direction, $N_x$.

\begin{figure}
    \centering

    \includegraphics[width=0.8\textwidth]{2D_mesh}

    \caption{Overview of the 2-D computational mesh.}

    \label{fig:2d-br-mesh}
\end{figure}

\begin{figure}
    \centering

    \includegraphics[width=0.7\textwidth]{2D_blade_mesh_closeup}

    \caption{Detailed view of the 2-D computational mesh near the blades.}

    \label{fig:blade-mesh}
\end{figure}

\subsection{Solver}

Simulations were run using OpenFOAM's pimpleDyMFoam solver, which uses a hybrid
PISO-SIMPLE algorithm for pressure-velocity coupling and is compatible with
dynamic meshes. An Euler scheme was used to advance the simulation forward in
time. The case files required to replicate the simulations are available from
\cite{Bachant2016-UNH-RVAT-2D-OpenFOAM-SST, Bachant2016-UNH-RVAT-2D-OpenFOAM-SA,
Bachant2016-UNH-RVAT-3D-OpenFOAM-SST, Bachant2016-UNH-RVAT-3D-OpenFOAM-SA}.

\subsection{Initial and boundary conditions}

Initial and boundary conditions were set to approximate those of the tow tank as
closely as possible. The velocity at the inlet, bottom, and side walls was fixed
to 1 m/s to match the tow tank case, while the top boundary condition was a slip
velocity condition. Pressure was held fixed at the outlet while a zero-gradient
boundary condition was applied at the inlet (a typical
velocity-inlet/pressure-outlet case). Note that in 2-D the top and bottom
boundary conditions are ``empty,'' which is an OpenFOAM convention to indicate
two-dimensionality.

Since the mesh was only refined next to the blade surfaces in order to resolve
the boundary layer profile, no wall functions were used. However, wall functions
were used for the other turbine surfaces, i.e., struts, hub, and shaft, and for
the tank sidewalls, bottom, and top. This strategy was adopted to reduce the
overall cell count; since the struts should be operating at nearly zero angle of
attack, it was assumed that separation would not be a signifigant contributor to
their parasitic drag.

\section{Model verification}

Both the $k$--$\omega$ SST and Spalart--Allmaras RANS model cases were verified
for convergence of the turbine mean power coefficient with respect to grid
spacing and time step using 2-D domains. The grid topology was fixed, but the
number of cells per domain length $N_x$ was scaled proportionally, maintaining
the same background mesh cell aspect ratio. Results for the $N_x$ parameter
sweep are shown in Figure~\ref{fig:2d-br-verification}, from which the final
number of streamwise grid points $N_x = 70$ was chosen, which resulted in a
nondimensional wall distance $y^+ = u^* y / \nu \sim 1$, where $u^*$ is the
friction velocity, defined as $u^*=\sqrt{\tau_w / \rho}$, where $\tau_w$ is the
wall shear stress. This mesh resolution gave a total cell count of approximately
$5 \times 10^4$ for the 2-D cases and 16 million for the 3-D case. Sensitivity
to domain length was assessed as well, showing a 2\% increase in mean power
coefficient for the 2-D SST case with the domain extended an additional two
rotor diameters downstream.

\begin{figure}
    \centering

    \includegraphics[width=0.9\textwidth]{verification}

    \caption{Time step (left) and grid size (right) dependence for the 2-D case
    with both the SST and SA turbulence models. Time step dependence was carried
    out with $N_x=70$ and grid size dependence with the time steps annotated for
    each turbulence model.}

    \label{fig:2d-br-verification}
\end{figure}

Time step dependence was evaluated using the 2-D $N_x=70$ grid, the results from
which are shown in Figure~\ref{fig:2d-br-verification}. It was seen that the
Spalart--Allmaras model converged well with decreasing time step, leading to a
final time step of 0.001 s. The results from the SST model show a local minimum
at $\Delta t = 0.002$ s, with some divergence for smaller time steps, the reason
for which is unknown. The local minimum was chosen as the final time step to run
the simulations. Note that the SST model's convergence behavior may be due to
its specific implementation in OpenFOAM, and not indicative of the nature of the
model equations. Note that while in general there may be a coupling between
necessary spatial and temporal resolution, these values were arrived at
iteratively, i.e., the spatial dependence was varied with the chosen time step,
and vice versa, such that there was one independent variable for each parameter
sweep. Reassuringly, the chosen time step and average $y^+$ adjacent to the
blades are comparable to others reported in the literature
\cite{Maitre2013,Balduzzi2016}.

Sensitivity to spatial resolution and time step was not assessed for the 3-D
case, as the computational cost would have been prohibitive. However, the mesh
resolution in the $x$--$y$ plane and time step were chosen based on the 2-D
results. Selecting a grid and time step in this way presents an alternative that
may be useful to practitioners who do not have access to the computationl
resources necessary to perform multiple high fidelity 3-D simulations.

\section{Results}

Turbine operation in all cases was simulated for 10 seconds, or approximately
six rotor revolutions, to allow the revolution-averaged performance coefficient
to converge to within $\sim 1$ percentage point; cf.
Figure~\ref{fig:cp-convergence}. Computations for the 3-D cases were run on 192
processes (24 nodes $\times$ 8 cores each on Sandia National Labs' Red Mesa high
performance computing cluster) and took on the order of 1,000 CPU hours per
second of simulated time. The 2-D simulations were run on a single processor and
took on the order of one CPU hour per simulated second.

\begin{figure}
    \centering

    \includegraphics[width=0.75\textwidth]{cp-time-series}

    \caption{Revolution-averaged power coefficient versus rotor revolution
    simulated with the 3-D $k$--$\omega$ SST model.}

    \label{fig:cp-convergence}
\end{figure}

Performance statistics were computed after the first revolution onward and flow
statistics were calculated over the time interval spanning $5$--$10$ seconds, or
approximately three rotor revolutions, from time series downsampled to 50 Hz. It
is assumed that the downsampling frequency is high enough above the blade
passage frequency such that differences from the variance (for computing
turbulence statistics) in the original velocity will be negligible.

\subsection{Performance prediction}

Predictions for both the mean rotor power coefficient $C_P$ and drag coefficient
\begin{equation}
    C_D = \frac{F_D}{\frac{1}{2} \rho A U_\infty^2},
    \label{eq:cd}
\end{equation}
where $F_D$ is the total streamwise force acting on the rotor, at the optimal
tip speed ratio $\lambda_0 = 1.9$ are shown in
Figure~\ref{fig:br-cfd-perf-bar-chart}. In general, the 2-D CFD cases both
significantly overpredict turbine loading and therefore mechanical power output,
which is due to their increased blockage ratio, unresolved blade end effects,
and lack of blade support struts. Effects of blade support strut drag were
investigated experimentally by Rawlings~\cite{Rawlings2008}, showing that for a
rotor with six foil-shaped struts, parasitic losses would cause a 4 percentage
point reduction in $C_P$ at $\lambda=2$. Since the UNH-RVAT has only three
foil-shaped struts, it therefore follows that the overprediction by the 2-D
simulations is mostly attributable to blockage and end effects, rather than
support strut drag.

The 3-D simulations matched the experimental measurements more closely. The
Spalart--Allmaras model's mean power coefficient was 0.24 while the SST's was
0.33, which represent a 6\% under- and 30\% overprediction, respectively,
compared with the experiments. Strut drag modeling discrepancies due to the
struts' lower mesh resolution compared to the blades, along with the use of wall
functions, could have played a minor role in degrading overall performance
predictions, and should be investigated further.

Some of the apparent overprediction of rotor drag coefficient could also be an
effect of the experimental procedure, where the ``tare drag'' from the turbine
mounting structure was measured without a turbine installed, then subtracted in
post-processing. With the turbine installed, flow past the mounting frame will
be higher due to blockage, meaning the tare drag would be underestimated, and
the CFD results for $C_D$ should be slightly higher than the experimental
measurements.

\begin{figure}
    \centering

    \includegraphics[width=0.95\textwidth]{perf_bar_chart}

    \caption{Power (left) and rotor drag (right) coefficient predictions at
    $\lambda = 1.9$, from experiments and each numerical model.}

    \label{fig:br-cfd-perf-bar-chart}
\end{figure}

\subsection{Wake characteristics}

Visualizations of the complex vorticity field generated by the turbine are
presented for the 2-D and 3-D Spalart--Allmaras cases in
Figure~\ref{fig:br-vorticity-2d} and Figure~\ref{fig:br-vorticity-3d},
respectively. It can be seen how the upstream blade---as it turns back into the
streamwise direction---is shedding a large amount of spanwise vorticity due to
the separated flow. In the 3-D case, strong tip vortices are also present, which
trace the ``contracting'' wake flow on the $-y$ side of the turbine associated
with the induced vertical velocity field. The 3-D dynamic stall vortex also
shows asymmetry about the $x$--$y$ mid-rotor plane; once again highlighting the
importance of three-dimensional effects on wake dynamics.

\begin{figure}
    \centering

    \includegraphics[width=0.75\textwidth]{2D_vorticity_SA_964}

    \caption{Instantaneous vorticity contours (at $t=9.64$ s) computed for the
        2-D Spalart--Allmaras case.}

    \label{fig:br-vorticity-2d}
\end{figure}

\begin{figure}
    \centering

    \includegraphics[width=0.75\textwidth]{3D_vorticity_SA_964_10-threshold}

    \caption{Iso-vorticity contours (at $t=9.64$ s) colored by the streamwise
        component of vorticity for the 3-D Spalart--Allmaras case.}

    \label{fig:br-vorticity-3d}
\end{figure}

Mean velocity profiles at the mid $x$--$y$ plane (or the nondimensional vertical
coordinate $z/H=0$) one turbine diameter downstream are shown in
Figure~\ref{fig:br-cfd-profiles}. The 2-D results suffer from a blockage
mismatch, i.e., keeping the proximity of the walls constant increases the
blockage ratio. The 3-D results, however, show good agreement with the
experiments.

\begin{figure}
    \centering

    \includegraphics[width=0.95\textwidth]{profiles}

    \caption{Mean velocity (left) and turbulence kinetic energy (right) profiles
    at $x/D=1$ from 2-D simulations, 3-D simulations (at the mid $x$--$y$ plane,
    or $z/H=0$), and experiments \cite{Bachant2015-JoT}.}

    \label{fig:br-cfd-profiles}
\end{figure}

Since this is a relatively high Reynolds number flow, it was of interest to
evaluate turbulence predictions, since these will play an important role in
overall flow dynamics. Validation was performed against the UNH-RVAT's
experimental dataset, which was acquired via acoustic Doppler velocimetry. The
turbulence kinetic energy for the unsteady RANS models was calculated as
\begin{equation}
    k = k_{\mathrm{RA}} + \frac{1}{2} \left(
    \overline{U^\prime}^2 +
    \overline{V^\prime}^2 +
    \overline{W^\prime}^2 \right),
    \label{eq:k}
\end{equation}
where $U^\prime = U - \overline{U}$ (the resolved velocity fluctuations) and
$k_{\mathrm{RA}}$ is the kinetic energy calculated by the turbulence model,
which is zero for the SA model. Profiles from the simulation results are shown
in Figure~\ref{fig:br-cfd-profiles}.

Both Spalart--Allmaras cases did a poor job predicting the turbulence kinetic
energy in the flow, since it must be resolved as variance in the velocity field.
The 2-D SST model did a good job predicting the peak in $k$ at $y/R=-1$, where
$y$ is the cross-stream coordinate and $R$ is the rotor radius, though is
missing the smaller peak at $y/R=-1$. This is once again likely due to blockage
issues, where local tip speed ratio is decreased, increasing the blades'
instantaneous angle of attack at this location on the downstream passage. In
contrast, the 3-D SST model predicts the $+y$ peak in turbulence kinetic energy
very well, though the $-y$ peak magnitude is overpredicted by about 30\%. We
also see some smearing of $k$ across the center of the rotor, which is likely
due to exaggerated levels of the turbulent eddy viscosity. This could also be an
effect of mesh resolution, though for this study only base mesh size sensitivity
was assessed, rather than local refinement, as this would have expanded the
parameter space significantly.

\subsubsection{Mean velocity in three dimensions}

In order to visualize the mean velocity field, vector arrows for the mean
cross-stream and vertical components are superimposed on top of contours of the
streamwise component at one rotor diameter downstream ($x/D=1$) of the axis in
Figure~\ref{fig:br-cfd-mean-velocity}. Both CFD models predict the general
structure of the mean velocity well, though the SA case has a slightly larger
vertical mean flow component, which could be due to stronger tip vortex
generation, or lower diffusivity compared with the SST model.

\begin{figure}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \centering

        \includegraphics[clip, trim=0 0.1in 0.3in 0.2in,
        width=0.72\textwidth]{meancontquiv_exp}

        \caption{Mean velocity field at $x/D=1$ from acoustic Doppler
        velocimetry measurements in a towing
        tank~\cite{Bachant2016-RVAT-Re-dep}.}

        \label{fig:br-cfd-meancontquiv-exp}
    \end{subfigure}

    \begin{subfigure}[b]{\textwidth}
        \centering

        \includegraphics[clip, trim=0 0.2in 0 0.15in,
        width=0.8\textwidth]{meancontquiv_kOmegaSST}

        \caption{Mean velocity at $x/D=1$ computed by the 3-D SST model.}

        \label{fig:meancontquiv-SST}
    \end{subfigure}

    \begin{subfigure}[b]{\textwidth}
        \centering

        \includegraphics[clip, trim=0 0.2in 0 0.15in,
        width=0.8\textwidth]{meancontquiv_SpalartAllmaras}

        \caption{Mean velocity at $x/D=1$ computed by the 3-D SA model.}

        \label{fig:meancontquiv-SA}
    \end{subfigure}

    \caption{Mean velocity from experiments and 3-D CFD cases. Solid gray lines
        indicate turbine frontal area and dashed lines indicate experimental
        measurement plane.}

    \label{fig:br-cfd-mean-velocity}
\end{figure}

\subsubsection{Turbulence kinetic energy contours}

Turbulence kinetic energy contours for the experimental measurements and each
CFD case at $x/D=1$ are presented in Figure~\ref{fig:br-cfd-kcont}. As seen in
the profiles in Figure~\ref{fig:br-cfd-profiles}, the SA model was not able to
resolve the majority of the unsteadiness in the flow. In contrast, the SST model
did a good job predicting the locations and magnitudes of various peaks in $k$.
These are generated along the top of the turbine via tip vortex shedding, and
the $-y$ side of the turbine via dynamic stall. We do however see the smearing
effect from the dynamic stall vortex centered around $z/H=0$, which is likely
more of an issue with wake evolution rather than wake generation. The effects of
mesh resolution on the breakdown and diffusion of the dynamic stall vortex may
also have an influence here.

\begin{figure}
    \centering
    \begin{subfigure}[b]{\textwidth}
        \centering

        \includegraphics[width=0.95\textwidth]{kcont_exp}

        \caption{Turbulence kinetic energy at $x/D=1$ from acoustic Doppler
        velocimetry measurements in a towing
        tank~\cite{Bachant2016-RVAT-Re-dep}.}

        \label{fig:kcont-exp}
    \end{subfigure}

    \begin{subfigure}[b]{\textwidth}
        \centering

        \includegraphics[width=0.95\textwidth]{kcont_kOmegaSST}

        \caption{Turbulence kinetic energy at $x/D=1$ computed by the 3-D SST
            model.}

        \label{fig:kcont-SST}
    \end{subfigure}

    \begin{subfigure}[b]{\textwidth}
        \centering

        \includegraphics[width=0.95\textwidth]{kcont_SpalartAllmaras}

        \caption{Turbulence kinetic energy at $x/D=1$ computed by the 3-D SA
            model.}

        \label{fig:kcont-SA}
    \end{subfigure}

    \caption{Turbulence kinetic energy from experiments and 3-D CFD cases. Solid
        black lines indicate turbine frontal area.}

    \label{fig:br-cfd-kcont}
\end{figure}

\subsubsection{Momentum recovery}

To get an overall idea of the wake recovery predicted by each model, and
therefore its potential effectivness in evaluating array spacing, the streamwise
component of the Navier--Stokes equation was rearranged to isolate the
streamwise partial derivative of the streamwise mean velocity $\partial U /
\partial x$---following Bachant and Wosnik~\cite{Bachant2015-JoT}---and compute
each term at $x/D = 1$ to compare with the experimental results.

We use the RANS models' eddy viscosity to calculate the turbulent transport via
\begin{equation}
    \text{Turb. trans.} = \nu_t \nabla^2 \vec{U},
    \label{eq:turb-trans}
\end{equation}
which is a different approach from those taken on the experiments, where
Reynolds stresses were measured, but $x$-derivatives were not:
\begin{equation}
    \text{Turb. trans. (exp.)} =
    -\left(
    \frac{\partial}{\partial y} \overline{u^\prime v^\prime}
    +
    \frac{\partial}{\partial z} \overline{u^\prime w^\prime}
    \right).
\end{equation}
As such, we should not be surprised if the CFD models predict higher levels of
turbulent transport than the experiments.

Normalized weighted averages for each recovery term at $x/D=1$ are computed and
multiplied by the cross-sectional area of the measurement plane, or the channel
width in the 2-D cases. Results are shown in a bar chart in
Figure~\ref{fig:br-cfd-recovery}. Consistent with the relatively large Reynolds
number regime, viscous transport is essentially negligible compared with other
mechanisms. Cross-stream advection $-V \frac{\partial U}{\partial y}$; or the
tendency of streamlines to diverge and reduce the streamwise momentum---produces
a negative effect for all cases, though the 3-D SST model predicts significantly
lower values. Mean vertical advection $-W \frac{\partial U}{\partial z}$ is by
definition zero for the 2-D cases. The 3-D cases show varying results---with the
SST model overpredicting and SA underpredicting the vertical velocity's effect
on replenishing streamwise momentum. Nevertheless, both 3-D models were able to
qualitatively resolve this mean vertical advection caused by blade tip vortex
shedding, which will be crucial to predicting overall wake recovery and the
performance of closely spaced turbines.

Turbulent transport and mean streamwise pressure gradient ($-\partial P \partial
x$) terms show the largest discrepancy between results. The 3-D SST case,
despite doing a good job predicting turbulence kinetic energy, significantly
overpredicts the turbulent transport term, while other CFD cases are comparable
with the experiment. This seems to be balanced by a large adverse pressure
gradient, which is also present to a smaller degree in the 3-D SA case.
Interestingly, in contrast, both 2-D CFD cases create a wake where the pressure
gradient is acting to slightly accelerate the flow at $x/D=1$. Unfortunately,
pressure data were not acquired from the experiment, though results from the 3-D
delayed detached eddy simulation (DDES) of Boudreau and Dumas
\cite{Boudreau2015} concur with the adverse pressure gradient condition.

\begin{figure}
    \centering

    \includegraphics[width=0.9\textwidth]{mom_bar_graph}

    \caption{Weighted sum normalized momentum recovery terms for each CFD case
        and experiments\cite{Bachant2016-RVAT-Re-dep} at $x/D=1$.}

    \label{fig:br-cfd-recovery}
\end{figure}

\section{Conclusions}

A cross-flow turbine was modeled using the $k$--$\omega$ SST and
Spalart--Allmaras (SA) Reynolds-averaged Navier--Stokes turbulence models to
test their abilities to predict turbine performance and near-wake dynamics. It
was observed that when modeled in 2-D, performance is overpredicted compared to
the data from the tow tank experiments, which was expected due to omission of
blade end effects, support strut drag, and increased blockage. Vertical (or
axial) wake dynamics were unresolved in the 2-D model, despite being identified
as a significant contributor to streamwise wake recovery, which casts doubt on
the 2-D model's applicability as a tool to study array spacing effects.

The 3-D blade-resolved RANS simulations predicted turbine performance and
near-wake quite well, where the SA model results were closest to the
experimentally measured performance. The SA model's effectiveness gives hope for
the prospect of replacing some physical modeling with blade-resolved CFD.
However, the  overprediction of power coefficient by the SST model highlights
the level of technological and economic risk involved with CFD. That is to say,
the SST model, despite being considered relatively trustworthy in these flow
conditions, would overestimate power output by 30\%.

Both the SST and SA models did a good job predicting the mean velocity field at
one turbine diameter downstream, but the SST model was more effective at
predicting turbulence kinetic energy, since it is solved for in the turbulence
model equations. Streamwise momentum recovery terms were computed from the CFD
results over an entire cross-section of the domain at $x/D=1$. Values for
transport due to the mean pressure gradient and turbulent fluctuations varied a
lot between the two turbulence models. Both models, however, were able to at
least qualitatively resolve the vertical component of the near-wake's mean
velocity field, which will be crucial to predicting the performance of closely
spaced CFTs.

The effectiveness at predicting the mean velocity field gives credibility to the
prospect of using the computed flow field to extrapolate the experimental
results, such that the CFD results can be used as a target for a low-order wake
generator or force parameterization. These results may also help develop new tip
loss corrections for blade element type models, which currently only exist for
axial-flow rotors, since they provide access to the pressure and shear forces
over the entire blade surface, which were not measured experimentally.
Ultimately, however, the computational cost of 3-D simulations---about 1,000 CPU
hours per simulated second (or roughly 10,000 CPU hours per operating
point)---may be too expensive to be used for CFT engineering work, especially
considering the uncertainty involved compared with physical model studies. Since
the 3-D wake results appeared to be only weakly asymmetrical in the $x$--$y$
plane about $z/H=0$, it may be possible to reduce computing power by only
modeling the top half of the rotor and using a symmetry boundary condition at
the mid-plane. However, 3-D blade-resolved RANS will likely not be practical for
turbine array simulations for quite some time, and until they are, low-order
models will need to be developed and improved for this purpose.

\begin{acknowledgments}
    The authors acknowledge funding through a National Science Foundation CAREER
    award (principal investigator Martin Wosnik, NSF 1150797, Energy for
    Sustainability, program manager Gregory L. Rorrer). The authors also thank
    Dr. Vincent S. Neary and the Sandia National Laboratories Water Power
    Program for use of their Red Mesa high performance computing cluster.
\end{acknowledgments}

% Create the reference section using BibTeX:
\bibliography{library}

\end{document}